Notes on Validated Model Counting \\ Version of \today

### 2023-03-09 doc/mc-notes.tex

\begin{table}
  \caption{CRAT Step Types.  $C$: clause identifier, $L$: literal, $V$: variable}
  \label{tab:crat:syntax}
\centering{
  \begin{tabular}{lllll}
    \multicolumn{4}{c}{Rule} & \multicolumn{1}{c}{Description} \\
    \midrule
    \makebox[5mm][l]{$C$} & \makebox[10mm][l]{\tt a}  & \makebox[15mm][l]{$L^{*}$ {\tt 0}} & \makebox[15mm][l]{$C^{+}$ {\tt 0}}  & \makebox[20mm][l]{Add RUP clause.  Explicit hint} \\
    $C$ & {\tt a} & $L^{*}$ {\tt 0} &  {\tt * 0} & Add RUP clause.  Inferred hint \\
     & {\tt dc} & $C$             & $C^{+}$  {\tt 0} & Delete RUP clause.  Explicit hint \\
     & {\tt dc} & $C$             & {\tt * 0} & Delete RUP clause.  Inferred hint \\
    \midrule
    $C$    & {\tt p} & $V \; L^{2+}$ {\tt 0}    &                  & Declare $\pand$ operation \\
    $C$    & {\tt s} & $V \; L \; L$    & $C^{+}$ {\tt 0}  & Declare $\por$ operation.  Explicit hint \\ 
    $C$    & {\tt s} & $V \; L \; L$    & {\tt * 0}  & Declare $\por$ operation.  Inferred hint \\ 
%%        & {\tt ite} & $V \; L \; L\; L$    &                  & Declare $\pite$ operation \\
     & {\tt do} & $V$             &  & Delete operation. \\

  \end{tabular}
  }
\end{table}

@@

The declaration of a $\pand$ operation has the form:
\begin{center}
\begin{tabular}{ccccccccc}
  \makebox[5mm]{$i$} & \makebox[5mm]{{\tt p}} & \makebox[5mm]{$v$} & \makebox[5mm]{$l_1$} & \makebox[5mm]{$l_2$} &
  \makebox[5mm]{$\cdots$} & \makebox[5mm]{$l_k$} & \makebox[5mm]{\tt 0} \\
\end{tabular}
\end{center}
where $i$ is a new clause ID, $v$ is a positive integer that does not
correspond to any previous variable, and $l_1, l_2, \ldots, l_k$ is a sequence of $k$  ($k \geq 2$)
integers representing literals of existing variables.  
This declaration implicitly causes $k+1$ clauses to be added to the proof:
\begin{center}
\begin{tabular}{cccccc}
\makebox[10mm]{ID} & \multicolumn{5}{c}{Clause} \\
  $i$ & $v$ & $-l_1$ & $-l_2$ & $\cdots$ & $-l_k$\\
  $i\!+\!1$ & $-v$ & $l_1$  \\
  $i\!+\!2$ & $-v$ & $l_2$  \\
  & $\ldots$ \\
  $i\!+\!k$ & $-v$ & $l_k$  \\
\end{tabular}
\end{center}
The dependency sets for the arguments represented by each pair of literals $l_{j_1}$ and $l_{j_2}$, for $1 \leq j_1 < j_2 \leq k$, must be disjoint.

@@

Finally, operator deletion ({\tt do}) specifies which operator to
delete via the extension variable associated with the operator.  Doing
so for operator $v$ requires that there is no
(undeleted) RUP clause containing a literal of $v$.
Implicitly, that means
that there is no (undeleted) operator $u$
having a literal of $v$ as one of its inputs.  The defining clauses for the operation are also deleted.

@@

A CRAT proof follows the same general form as a QRAT dual
proof~\cite{bryant:cade:2021}---one that ensures that each clause
addition and each clause deletion preserves equivalence.  With CRAT,
however, clauses are defined both explicitly and implicitly.  Starting
with the set of input clauses, the proof consists of a sequence of steps that
both add and delete clauses.  Each addition must be truth preserving
and each deletion must be falsehood preserving.  At the end, the following conditions must hold:
\begin{itemize}
\item All of the input clauses should have been deleted.
\item There should be exactly one clause left that was added via RUP addition, and this should be a unit clause.
\end{itemize}
The remaining
unit clause consists of some variable or its
complement.  Except for trivial cases, the final literal will be an
extension variable or its complement.  That literal will indicate the
root of the POG as the (possibly negated) output of a POG
operation.  Working from that root backward, the POG can be
extracted from the CRAT file.

@@

\begin{figure}
  A).  CRAT File Contents
  \begin{center}
  \begin{tabular}{lcllll}
    \multicolumn{4}{c}{File line} & & \multicolumn{1}{c}{Explanation} \\
\midrule
    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{\tt i}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause}\\
    {\tt 2}        & {\tt p}   & {\tt 4 -2 3 0}  &            & & Declare $p_4 = \obar{x}_2 \pand x_3$ \\
    {\tt 5}        & {\tt s}   & {\tt 5 2 4}   & {\tt 3 0}  & & Declare $s_5 = x_2 \por p_4$ \\
    {\tt 8}        & {\tt p}   & {\tt 6 -1 5 0}  &            & & Declare $p_6 = \obar{x}_1 \pand s_5$ \\
    {\tt 11}         & {\tt s}   & {\tt 7 1 6}   & {\tt 9 0}  & & Declare $s_7 = x_1 \por p_6$ \\
    {\tt 14} & {\tt a}  & {\tt 7 0} & {\tt 13 12 8 7 6 2 1 0} & & Assert unit clause $[s_7]$ \\
             & {\tt dc}  & {\tt 1}  & {\tt 4 5 10 11 14 0} & & Delete input clause \\
  \end{tabular}
  \end{center}
B). Proof Clauses
    \begin{center}
  \begin{tabular}{lcllll}
    \multicolumn{4}{c}{Clause} & & \multicolumn{1}{c}{Explanation} \\
\midrule
    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause}\\
    {\tt 2} &     & {\tt 4  2 -3 0}  &     & & Defining clauses for $p_4$ \\ 
    {\tt 3} &     & {\tt -4 -2 0}    & & & \\ %%Defining clause for $p_4$ \\ 
    {\tt 4} &     & {\tt -4 3 0}    &  & & \\ %%Defining clause for $p_4$ \\ 
            &     & {\tt -2 -4 0}   & {\tt 3 0}  & &  Mutual exclusion proof for $s_5$ \\
    {\tt 5} &     & {\tt -5 2 4 0}  &     & & Defining clauses for $s_5$ \\ 
    {\tt 6} &     & {\tt  5 -2 0}    &  &  \\  %%Defining clause for $s_5$ \\ 
    {\tt 7} &     & {\tt  5 -4 0}    &  & &  \\  %%Defining clause for $s_5$ \\ 
    {\tt 8} &     & {\tt 6  1 -5 0}  &    & & Defining clauses for $p_6$ \\ 
    {\tt 9} &     & {\tt -6 -1 0}    &  & &  \\  %%Defining clause for $p_6$ \\ 
    {\tt 10} &     & {\tt -6 5 0}    &  & &  \\  %%Defining clause for $p_6$ \\ 
            &     & {\tt -1 -6 0}   & {\tt 9 0}  & &  Mutual exclusion proof for $s_7$ \\
    {\tt 11} &     & {\tt -7 1 6 0}  &  & & Defining clauses for $s_7$ \\ 
    {\tt 12} &     & {\tt  7 -1 0}     & &  \\  %%Defining clause for $s_7$ \\ 
    {\tt 13} &     & {\tt  7 -6 0}     & &  \\  %%Defining clause for $s_7$ \\ 
    {\tt 14} &     & {\tt 7 0} & {\tt 13 12 7 6 2 1 0} & & Assert unit clause $[s_7]$ \\
             & {\tt dc}  & {\tt 1}  & {\tt 4 5 10 11 14 0} & & Delete input clause \\
  \end{tabular}
  \end{center}
  \caption{CRAT file \#1 for formula $x_1 \lor x_2 \lor x_3$, and the resulting set of proof clauses}
  \label{fig:c3:crat}
\end{figure}

@@

    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{\tt i}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause $C_1$}\\
    {\tt 2} & {\tt i}   & {\tt -1 2 0}  &            & & Input clause $C_2$ \\
    {\tt 3} & {\tt p}   & {\tt 4 -1 -2 -3 0}  &            & & Declare $p_4 = \obar{x}_1 \pand \obar{x}_2 \pand \obar{x}_3$ \\
    {\tt 7} & {\tt a}  & {\tt -4 0} & {\tt 6 5 4 1 0} & & Assert $t_1 = \obar{p}_4 = x_1 \lor x_2 \lor x_3$ \\
            & {\tt dc}  & {\tt 1 } & {\tt 3 7 0} & & Delete clause $C_1$\\
    {\tt 8} & {\tt p}   & {\tt 5 1 -2 0} &              & & Declare $p_5 = x_1 \pand \obar{x}_2$ \\    
    {\tt 11} & {\tt a} & {\tt -5 0} & {\tt 10 9 2 0}  & & Assert $t_2 = \obar{p}_5 = \obar{x}_1 \lor x_2$ \\
            & {\tt dc}  & {\tt 2 } & {\tt 8 11 0} & & Delete clause $C_2$\\
    {\tt 12} & {\tt s}   & {\tt 6 4 5}   & {\tt 4 9 0}  & & Declare $s_6 = \obar{t}_1 \por \obar{t}_2$ \\
    {\tt 15} & {\tt a}  & {\tt -6 0}       & {\tt 7 11 12 0} & & Assert $t_3 = \obar{s}_6 = t_1 \land t_2$ \\
             & {\tt dc}  & {\tt 7}          & {\tt 13 15 0} & & Delete $t_1$\\
             & {\tt dc}  & {\tt 11}         & {\tt 14 15 0} & & Delete $t_2$\\
  \end{tabular}
  \end{center}  
B). Proof Clauses
  \begin{center}
  \begin{tabular}{lcllll}
    \multicolumn{4}{c}{Clause} & & \multicolumn{1}{c}{Explanation} \\
\midrule
    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause $C_1$}\\
    {\tt 2} &    & {\tt -1 2 0}  &            & & Input clause $C_2$ \\
    {\tt 3} &   & {\tt 4 1 2 3 0} &     & & Defining clauses for $p_4$ \\
    {\tt 4} &   & {\tt -4 -1 0} &     & & \\
    {\tt 5} &   & {\tt -4 -2 0} &     & & \\
    {\tt 6} &   & {\tt -4 -3 0} &     & & \\
    {\tt 7} &   & {\tt -4 0} & {\tt 6 5 4 1 0} & & Assert $t_1 = \obar{p}_4 = x_1 \lor x_2 \lor x_3$ \\
            & {\tt dc}  & {\tt 1 } & {\tt 3 7 0} & & Delete clause $C_1$\\

@@

\begin{figure}
  A).  CRAT File Contents
  \begin{center}
  \begin{tabular}{llllll}
    \multicolumn{4}{c}{File line} & & \multicolumn{1}{c}{Explanation} \\
\midrule
    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause $C_1$}\\
            {\tt 2} & {\tt i}   & {\tt -1 2 0}  &            & & Input clause $C_2$ \\
            {\tt 3} & {\tt p}   & {\tt 4 -2 -3 0}  &            & & Declare $p_4 = \obar{x}_2 \pand \obar{x}_3$ \\
    {\tt 6} & {\tt p}   & {\tt 5 -1 -4 0}  &             & & Declare $p_5 = \obar{x}_1 \pand \obar{p}_4$ \\
    {\tt 9} & {\tt p}   & {\tt 6 1 2 0} &              & & Declare $p_6 = x_1 \pand x_2$ \\    
    {\tt 12} & {\tt s}   & {\tt 7 5 6}   & {\tt 7 10 0}  & & Declare $s_7 = p_5 \por p_6$ \\
    {\tt 15} & {\tt a}  & {\tt 1 -4 0}    & {\tt 4 5 1 0} & & Justify $\obar{x}_1 \rightarrow \obar{p}_4$ \\
    {\tt 16} & {\tt a}  & {\tt 1 7 0}     & {\tt 13 6 15 0} & & Justify $\obar{x}_1 \rightarrow s_7$ \\
    {\tt 17} & {\tt a}  & {\tt -1 7 0}    & {\tt 14 9 2 0} & & Justify $x_1 \rightarrow s_7$ \\
    {\tt 18} & {\tt a}  & {\tt 7 0}       & {\tt 16 17 0}  & & Justify unit clause $t = s_7$ \\
             & {\tt dc}  & {\tt 1}         & {\tt 3 8 10 12 18 0} & & Delete $C_1$\\
             & {\tt dc}  & {\tt 2}         & {\tt 11 7 12 18 0} & & Delete $C_2$\\
  \end{tabular}
  \end{center}  
B). Proof Clauses
  \begin{center}
  \begin{tabular}{llllll}
    \multicolumn{4}{c}{File line} & & \multicolumn{1}{c}{Explanation} \\
\midrule
    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause $C_1$}\\
    {\tt 2} &    & {\tt -1 2 0}  &            & & Input clause $C_2$ \\
    {\tt 3} &   & {\tt 4 2 3 0} &     & & Defining clauses for $p_4$ \\
    {\tt 4} &   & {\tt -4 -2 0} &     & & \\
    {\tt 5} &   & {\tt -4 -3 0} &     & & \\
    {\tt 6} &   & {\tt 5 1 4 0} &     & & Defining clauses for $p_5$ \\
    {\tt 7} &   & {\tt -5 -1 0} &     & & \\
    {\tt 8} &   & {\tt -5 -4 0} &     & & \\
    {\tt 9} &   & {\tt 6 -1 -2 0}&     & & Defining clauses for $p_6$ \\ 
    {\tt 10} &   & {\tt -6 1 0}    &  & & \\
    {\tt 11} &   & {\tt -6 2 0}    &  & & \\ 
     &           & {\tt -5 -6}   & {\tt 7 10 0}  & & Mutual exclusion proof for $s_7$ \\
    {\tt 12} &   & {\tt -7 5 6 0}  &     & & Defining clauses for $s_7$ \\ 
    {\tt 13} &   & {\tt  7 -5 0}    &  & & \\  
    {\tt 14} &   & {\tt  7 -6 0}    &  & & \\
    {\tt 15} &   & {\tt 1 -4 0}    & {\tt 4 5 1 0} & & Justify $\obar{x}_1 \rightarrow \obar{p}_4$ \\
    {\tt 16} &   & {\tt 1 7 0}     & {\tt 13 6 15 0} & & Justify $\obar{x}_1 \rightarrow s_7$ \\
    {\tt 17} &   & {\tt -1 7 0}    & {\tt 14 9 2 0} & & Justify $x_1 \rightarrow s_7$ \\
    {\tt 18} &   & {\tt 7 0}       & {\tt 16 17 0}  & & Justify unit clause $t = s_7$ \\
             & {\tt dc}  & {\tt 1}         & {\tt 3 15 0} & & Delete $C_1$\\
             & {\tt dc}  & {\tt 2}         & {\tt 11 7 12 18 0} & & Delete $C_2$\\
  \end{tabular}
  \end{center}  
  \caption{CRAT file \#2B for formula $\phi_2 = (x_1 \lor x_2 \lor x_3) \land (\obar{x}_1 \lor x_2)$, and the resulting set of proof clauses}
  \label{fig:p2-cdcl:crat}
\end{figure}

>>>>

\begin{table}
  \caption{CRAT Step Types.  $C$: clause identifier, $L$: literal, $V$: variable}
  \label{tab:crat:syntax}
\centering{
  \begin{tabular}{lllll}
    \multicolumn{4}{c}{Rule} & \multicolumn{1}{c}{Description} \\
    \midrule
    \makebox[5mm][l]{$C$} & \makebox[10mm][l]{\tt a}  & \makebox[15mm][l]{$L^{*}$ {\tt 0}} & \makebox[15mm][l]{$C^{+}$ {\tt 0}}  & \makebox[20mm][l]{Add RUP clause.  Explicit hint} \\
    $C$ & {\tt a} & $L^{*}$ {\tt 0} &  {\tt * 0} & Add RUP clause.  Inferred hint \\
     & {\tt dc} & $C$             & $C^{+}$  {\tt 0} & Delete RUP clause.  Explicit hint \\
     & {\tt dc} & $C$             & {\tt * 0} & Delete RUP clause.  Inferred hint \\
    \midrule
    $C$    & {\tt p} & $V \; L^{*}$ {\tt 0}    &                  & Declare $\pand$ operation \\
    $C$    & {\tt s} & $V \; L \; L$    & $C^{+}$ {\tt 0}  & Declare $\por$ operation.  Explicit hint \\ 
    $C$    & {\tt s} & $V \; L \; L$    & {\tt * 0}  & Declare $\por$ operation.  Inferred hint \\ 
%%        & {\tt ite} & $V \; L \; L\; L$    &                  & Declare $\pite$ operation \\
     & {\tt do} & $V$             &  & Delete operation. \\
    \midrule
     & {\tt r} & $L$             &            & Declare root literal\\
  \end{tabular}
  }
\end{table}

@@

The declaration of a $\pand$ operation has the form:
\begin{center}
\begin{tabular}{ccccccccc}
  \makebox[5mm]{$i$} & \makebox[5mm]{{\tt p}} & \makebox[5mm]{$v$} & \makebox[5mm]{$l_1$} & \makebox[5mm]{$l_2$} &
  \makebox[5mm]{$\cdots$} & \makebox[5mm]{$l_k$} & \makebox[5mm]{\tt 0} \\
\end{tabular}
\end{center}
where $i$ is a new clause ID, $v$ is a positive integer that does not
correspond to any previous variable, and $l_1, l_2, \ldots, l_k$ is a sequence of $k$  ($k \geq 2$)
integers representing literals of existing variables.  
This declaration implicitly causes $k+1$ clauses to be added to the proof:
\begin{center}
\begin{tabular}{cccccc}
\makebox[10mm]{ID} & \multicolumn{5}{c}{Clause} \\
  $i$ & $v$ & $-l_1$ & $-l_2$ & $\cdots$ & $-l_k$\\
  $i\!+\!1$ & $-v$ & $l_1$  \\
  $i\!+\!2$ & $-v$ & $l_2$  \\
  & $\ldots$ \\
  $i\!+\!k$ & $-v$ & $l_k$  \\
\end{tabular}
\end{center}
The dependency sets for the arguments represented by each pair of
literals $l_{j_1}$ and $l_{j_2}$, for $1 \leq j_1 < j_2 \leq k$, must
be disjoint.  Note that a $\pand$ operation can have no arguments,
representing $\tautology$.  The only clause added to the proof will be
the unit literal $v$.  A reference to literal $-v$ then becomes a way
of representing $\nil$.

@@

Operator deletion ({\tt do}) specifies which operator to
delete via the extension variable associated with the operator.  Doing
so for operator $v$ requires that there is no
(undeleted) RUP clause containing a literal of $v$.
Implicitly, that means
that there is no (undeleted) operator $u$
having a literal of $v$ as one of its inputs.  The defining clauses for the operation are also deleted.

Finally, the literal denoting the root of the POG is declared with the
{\tt r} command.  It can occur anywhere in the file.  Typically, it
will be the extension node representing the root of a graph, but in
degenerate cases it can be the positive or negative literal of one of
the input variables.

@@

A CRAT proof follows the same general form as a QRAT dual
proof~\cite{bryant:cade:2021}---one that ensures that each clause
addition and each clause deletion preserves equivalence.  With CRAT,
however, clauses are defined both explicitly and implicitly.  Starting
with the set of input clauses, the proof consists of a sequence of steps that
both add and delete clauses.  Each addition must be truth preserving
and each deletion must be falsehood preserving.  At the end, the following conditions must hold:
\begin{itemize}
\item All of the input clauses should have been deleted.
\item There should be exactly one clause left that was added via RUP addition, and this should be a unit clause.
\end{itemize}
The remaining
unit clause consists of some variable or its
complement.  Except for trivial cases, the final literal will be an
extension variable or its complement.  That literal must match the one declared as the root literal.

@@

\begin{figure}
  A).  CRAT File Contents
  \begin{center}
  \begin{tabular}{lcllll}
    \multicolumn{4}{c}{File line} & & \multicolumn{1}{c}{Explanation} \\
\midrule
    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{\tt i}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause}\\
    {\tt 2}        & {\tt p}   & {\tt 4 -2 3 0}  &            & & Declare $p_4 = \obar{x}_2 \pand x_3$ \\
    {\tt 5}        & {\tt s}   & {\tt 5 2 4}   & {\tt 3 0}  & & Declare $s_5 = x_2 \por p_4$ \\
    {\tt 8}        & {\tt p}   & {\tt 6 -1 5 0}  &            & & Declare $p_6 = \obar{x}_1 \pand s_5$ \\
    {\tt 11}         & {\tt s}   & {\tt 7 1 6}   & {\tt 9 0}  & & Declare $s_7 = x_1 \por p_6$ \\
             & {\tt r}  & {\tt 7} & && Declare root literal \\
    {\tt 14} & {\tt a}  & {\tt 7 0} & {\tt 13 12 8 7 6 2 1 0} & & Assert unit clause $[s_7]$ \\
             & {\tt dc}  & {\tt 1}  & {\tt 4 5 10 11 14 0} & & Delete input clause \\
  \end{tabular}
  \end{center}
B). Proof Clauses
    \begin{center}
  \begin{tabular}{lcllll}
    \multicolumn{4}{c}{Clause} & & \multicolumn{1}{c}{Explanation} \\
\midrule
    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause}\\
    {\tt 2} &     & {\tt 4  2 -3 0}  &     & & Defining clauses for $p_4$ \\ 
    {\tt 3} &     & {\tt -4 -2 0}    & & & \\ %%Defining clause for $p_4$ \\ 
    {\tt 4} &     & {\tt -4 3 0}    &  & & \\ %%Defining clause for $p_4$ \\ 
            &     & {\tt -2 -4 0}   & {\tt 3 0}  & &  Mutual exclusion proof for $s_5$ \\
    {\tt 5} &     & {\tt -5 2 4 0}  &     & & Defining clauses for $s_5$ \\ 
    {\tt 6} &     & {\tt  5 -2 0}    &  &  \\  %%Defining clause for $s_5$ \\ 
    {\tt 7} &     & {\tt  5 -4 0}    &  & &  \\  %%Defining clause for $s_5$ \\ 
    {\tt 8} &     & {\tt 6  1 -5 0}  &    & & Defining clauses for $p_6$ \\ 
    {\tt 9} &     & {\tt -6 -1 0}    &  & &  \\  %%Defining clause for $p_6$ \\ 
    {\tt 10} &     & {\tt -6 5 0}    &  & &  \\  %%Defining clause for $p_6$ \\ 
            &     & {\tt -1 -6 0}   & {\tt 9 0}  & &  Mutual exclusion proof for $s_7$ \\
    {\tt 11} &     & {\tt -7 1 6 0}  &  & & Defining clauses for $s_7$ \\ 
    {\tt 12} &     & {\tt  7 -1 0}     & &  \\  %%Defining clause for $s_7$ \\ 
    {\tt 13} &     & {\tt  7 -6 0}     & &  \\  %%Defining clause for $s_7$ \\ 
    {\tt 14} &     & {\tt 7 0} & {\tt 13 12 7 6 2 1 0} & & Assert unit clause $[s_7]$ \\
             & {\tt dc}  & {\tt 1}  & {\tt 4 5 10 11 14 0} & & Delete input clause \\
  \end{tabular}
  \end{center}
  \caption{CRAT file \#1 for formula $x_1 \lor x_2 \lor x_3$, and the resulting set of proof clauses}
  \label{fig:c3:crat}
\end{figure}

@@

    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{\tt i}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause $C_1$}\\
    {\tt 2} & {\tt i}   & {\tt -1 2 0}  &            & & Input clause $C_2$ \\
    {\tt 3} & {\tt p}   & {\tt 4 -1 -2 -3 0}  &            & & Declare $p_4 = \obar{x}_1 \pand \obar{x}_2 \pand \obar{x}_3$ \\
    {\tt 7} & {\tt a}  & {\tt -4 0} & {\tt 6 5 4 1 0} & & Assert $t_1 = \obar{p}_4 = x_1 \lor x_2 \lor x_3$ \\
            & {\tt dc}  & {\tt 1 } & {\tt 3 7 0} & & Delete clause $C_1$\\
    {\tt 8} & {\tt p}   & {\tt 5 1 -2 0} &              & & Declare $p_5 = x_1 \pand \obar{x}_2$ \\    
    {\tt 11} & {\tt a} & {\tt -5 0} & {\tt 10 9 2 0}  & & Assert $t_2 = \obar{p}_5 = \obar{x}_1 \lor x_2$ \\
            & {\tt dc}  & {\tt 2 } & {\tt 8 11 0} & & Delete clause $C_2$\\
    {\tt 12} & {\tt s}   & {\tt 6 4 5}   & {\tt 4 9 0}  & & Declare $s_6 = \obar{t}_1 \por \obar{t}_2$ \\
             & {\tt r}  & {\tt -6} & && Declare root literal \\
    {\tt 15} & {\tt a}  & {\tt -6 0}       & {\tt 7 11 12 0} & & Assert $t_3 = \obar{s}_6 = t_1 \land t_2$ \\
             & {\tt dc}  & {\tt 7}          & {\tt 13 15 0} & & Delete $t_1$\\
             & {\tt dc}  & {\tt 11}         & {\tt 14 15 0} & & Delete $t_2$\\
  \end{tabular}
  \end{center}  
B). Proof Clauses
  \begin{center}
  \begin{tabular}{lcllll}
    \multicolumn{4}{c}{Clause} & & \multicolumn{1}{c}{Explanation} \\
\midrule
    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause $C_1$}\\
    {\tt 2} &    & {\tt -1 2 0}  &            & & Input clause $C_2$ \\
    {\tt 3} &   & {\tt 4 1 2 3 0} &     & & Defining clauses for $p_4$ \\
    {\tt 4} &   & {\tt -4 -1 0} &     & & \\
    {\tt 5} &   & {\tt -4 -2 0} &     & & \\
    {\tt 6} &   & {\tt -4 -3 0} &     & & \\
    {\tt 7} &   & {\tt -4 0} & {\tt 6 5 4 1 0} & & Assert $t_1 = \obar{p}_4 = x_1 \lor x_2 \lor x_3$ \\
            & {\tt dc}  & {\tt 1 } & {\tt 3 7 0} & & Delete clause $C_1$\\

@@

\begin{figure}
  A).  CRAT File Contents
  \begin{center}
  \begin{tabular}{llllll}
    \multicolumn{4}{c}{File line} & & \multicolumn{1}{c}{Explanation} \\
\midrule
    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause $C_1$}\\
            {\tt 2} & {\tt i}   & {\tt -1 2 0}  &            & & Input clause $C_2$ \\
            {\tt 3} & {\tt p}   & {\tt 4 -2 -3 0}  &            & & Declare $p_4 = \obar{x}_2 \pand \obar{x}_3$ \\
    {\tt 6} & {\tt p}   & {\tt 5 -1 -4 0}  &             & & Declare $p_5 = \obar{x}_1 \pand \obar{p}_4$ \\
    {\tt 9} & {\tt p}   & {\tt 6 1 2 0} &              & & Declare $p_6 = x_1 \pand x_2$ \\    
    {\tt 12} & {\tt s}   & {\tt 7 5 6}   & {\tt 7 10 0}  & & Declare $s_7 = p_5 \por p_6$ \\
             & {\tt r}  & {\tt 7} & && Declare root literal \\
    {\tt 15} & {\tt a}  & {\tt 1 -4 0}    & {\tt 4 5 1 0} & & Justify $\obar{x}_1 \rightarrow \obar{p}_4$ \\
    {\tt 16} & {\tt a}  & {\tt 1 7 0}     & {\tt 13 6 15 0} & & Justify $\obar{x}_1 \rightarrow s_7$ \\
    {\tt 17} & {\tt a}  & {\tt -1 7 0}    & {\tt 14 9 2 0} & & Justify $x_1 \rightarrow s_7$ \\
    {\tt 18} & {\tt a}  & {\tt 7 0}       & {\tt 16 17 0}  & & Justify unit clause $t = s_7$ \\
             & {\tt dc}  & {\tt 1}         & {\tt 3 8 10 12 18 0} & & Delete $C_1$\\
             & {\tt dc}  & {\tt 2}         & {\tt 11 7 12 18 0} & & Delete $C_2$\\
  \end{tabular}
  \end{center}  
B). Proof Clauses
  \begin{center}
  \begin{tabular}{llllll}
    \multicolumn{4}{c}{File line} & & \multicolumn{1}{c}{Explanation} \\
\midrule
    \makebox[5mm][l]{\tt 1} & \makebox[7mm]{}   & \makebox[20mm][l]{\tt 1 2 3 0}   &  \makebox[30mm]{}          & \makebox[5mm]{} & \makebox[40mm][l]{Input clause $C_1$}\\
    {\tt 2} &    & {\tt -1 2 0}  &            & & Input clause $C_2$ \\
    {\tt 3} &   & {\tt 4 2 3 0} &     & & Defining clauses for $p_4$ \\
    {\tt 4} &   & {\tt -4 -2 0} &     & & \\
    {\tt 5} &   & {\tt -4 -3 0} &     & & \\
    {\tt 6} &   & {\tt 5 1 4 0} &     & & Defining clauses for $p_5$ \\
    {\tt 7} &   & {\tt -5 -1 0} &     & & \\
    {\tt 8} &   & {\tt -5 -4 0} &     & & \\
    {\tt 9} &   & {\tt 6 -1 -2 0}&     & & Defining clauses for $p_6$ \\ 
    {\tt 10} &   & {\tt -6 1 0}    &  & & \\
    {\tt 11} &   & {\tt -6 2 0}    &  & & \\ 
     &           & {\tt -5 -6}   & {\tt 7 10 0}  & & Mutual exclusion proof for $s_7$ \\
    {\tt 12} &   & {\tt -7 5 6 0}  &     & & Defining clauses for $s_7$ \\ 
    {\tt 13} &   & {\tt  7 -5 0}    &  & & \\  
    {\tt 14} &   & {\tt  7 -6 0}    &  & & \\
    {\tt 15} &   & {\tt 1 -4 0}    & {\tt 4 5 1 0} & & Justify $\obar{x}_1 \rightarrow \obar{p}_4$ \\
    {\tt 16} &   & {\tt 1 7 0}     & {\tt 13 6 15 0} & & Justify $\obar{x}_1 \rightarrow s_7$ \\
    {\tt 17} &   & {\tt -1 7 0}    & {\tt 14 9 2 0} & & Justify $x_1 \rightarrow s_7$ \\
    {\tt 18} &   & {\tt 7 0}       & {\tt 16 17 0}  & & Justify unit clause $t = s_7$ \\
             & {\tt dc}  & {\tt 1}         & {\tt 3 15 0} & & Delete $C_1$\\
             & {\tt dc}  & {\tt 2}         & {\tt 11 7 12 18 0} & & Delete $C_2$\\
  \end{tabular}
  \end{center}  
  \caption{CRAT file \#2B for formula $\phi_2 = (x_1 \lor x_2 \lor x_3) \land (\obar{x}_1 \lor x_2)$, and the resulting set of proof clauses}
  \label{fig:p2-cdcl:crat}
\end{figure}
